---
title: "Colors"
author: "Ervin Madaha, Jennifer Wilken, Shuaib Yunus"
date: "5/21/2018"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=6, fig.height=6, fig.align="center")
```

## Setup

Load and format the data in a reusable way:
```{r load}
masterCard<-read.csv2("MasterColorCard.csv")

raw<-read.csv2("LabMeasurements-Color-Card.csv")

sheets<-rep(1:13,42)
raw$sheet<-sheets

column_names<-names(raw)[names(raw)!=c('Row','Column','sheet')]
l_data<-reshape(raw,varying = column_names, direction= "long", sep="" )
l_data$Xc<-floor(l_data$time/10)
l_data$Yc<-l_data$time-10*l_data$Xc
l_data$time<-NULL

l_data$X<-max(l_data$Xc)*(l_data$Row-1)+l_data$Xc
l_data$Y<-max(l_data$Yc)*(l_data$Column-1)+l_data$Yc

#Join the 2 data frames based on color coordinates (8x8)
joined_d<-merge(l_data,
                masterCard,
                by.x=c('Xc','Yc'),
                by.y=c('Crow','Ccol'),suffixes = c("_measured","_master"))
#center lab on master
joined_d$L_centered<-joined_d$L_measured-joined_d$L_master
joined_d$a_centered<-joined_d$a_measured-joined_d$a_master
joined_d$b_centered<-joined_d$b_measured-joined_d$b_master

#Compute sd based on center
L_sd<-sqrt(sum(joined_d$L_centered**2)/dim(joined_d)[1])
a_sd<-sqrt(sum(joined_d$a_centered**2)/dim(joined_d)[1])
b_sd<-sqrt(sum(joined_d$b_centered**2)/dim(joined_d)[1])
joined_d$L_scaled<-joined_d$L_centered/L_sd
joined_d$a_scaled<-joined_d$a_centered/a_sd
joined_d$b_scaled<-joined_d$b_centered/b_sd

#Most likely useless for pca
joined_d$X_scaled<-(joined_d$X-mean(joined_d$X))/sd(joined_d$X)
joined_d$Y_scaled<-(joined_d$Y-mean(joined_d$Y))/sd(joined_d$Y)
joined_d$sheet_scaled<-(joined_d$sheet-mean(joined_d$sheet))/sd(joined_d$sheet)
```

## Delta E computation

Load the library we'll be using:
```{r deltae-example, message=FALSE}
library("colorscience")
```

Let's create a convenience function for getting a Delta E (DE) vector from 2 L,a,b dataframes:
```{r deltae-helper}
getDEvector <-function(x, y){
  res <- 1:dim(x)[1]
  for (i in 1:dim(x)[1]){
    res[i] <- deltaE2000(data.matrix(x[i,]),data.matrix(y[i,]))
  }
  return(res)
}
```

Let's add a column for Delta E values from L, a, b values of measured versus master to our dataframe:
```{r add-de-col}
joined_d$DE <-
  getDEvector(joined_d[, c("L_measured", "a_measured", "b_measured")],
              joined_d[, c("L_master", "a_master", "b_master")])
```

Center and scale the Delta E column:
```{r center-delta-e}
DE_sd<-sqrt(sum((joined_d$DE-mean(joined_d$DE))**2)/dim(joined_d)[1])
joined_d$DE_scaled <- (joined_d$DE -mean(joined_d$DE))/DE_sd
```

Create a 64x546 matrix of DE values to recreate the color card DEM (DE Matrix).
Each row is a color on the 8x8 color card and columns are each color card (13x42(7x6))
```{r create-de-matrix}
DEM <-  matrix(joined_d$DE,64, byrow = TRUE)
DEM_mean <- rowMeans(DEM) # mean of each row to reduce each color to a single value
DEM_sd <- apply(DEM,1,sd) # Interesting to see how each color varies
```

Create an 8x8 matrix representing a card:
```{r}
DEM_avgcard <- matrix(DEM_mean,8,byrow = TRUE)
DEM_sd_scaledcard <- matrix(DEM_sd/max(DEM_sd),8,byrow = TRUE)
```

## Plots

One way to spread the color to express the DE spread across the color card:
```{r}
colrange <- colorRampPalette(c("yellowgreen","yellow","red"))(100)
```

Another way to express the color based on the relative human distinction:
```{r}
getDEcolor <-function(x){
  if (x < 1) "grey" # Unnoticable to the human eye
  else if (1 < x && x < 2) "yellowgreen" # barely noticable
  else if (2 < x && x < 3) "yellow" # noticable difference on a keen eye
  else if  (3 < x && x < 10) "orange" # distinguishable 
  else if (10 < x ) "red" # clearly distinct
}
```

Plot DE values with the 2nd color scheme only half:
```{r}
xspace<-cbind(1:100,1:100)
plot(xspace,type="n",axes=FALSE,xlab = "",ylab = "")
start<-5
width<-8
spacing<-2

for( i in 1:8){
  for( j in 1:8){
    xleft<-i*(width+spacing)
    ybottom<-j*(width+spacing)
    rect(xleft,ybottom,xleft+width,ybottom+width,col=getDEcolor(DEM_avgcard[i,j]),
         border = FALSE)
  }
}
```

Plot the right half of colors with sd following color scheme 1:
```{r}
xspace<-cbind(1:100,1:100)
plot(xspace,type="n",axes=FALSE,xlab = "",ylab = "")
start<-5
width<-10
spacing<-1

for( i in 1:8){
  for( j in 1:8){
    xleft<-i*(width+spacing)
    ybottom<-j*(width+spacing)
    rect(xleft,ybottom,xleft+width,ybottom+width,col=getDEcolor(DEM_avgcard[i,j]),
         border = FALSE)
  }
}

for( i in 1:8){
  for( j in 1:8){
    xleft<-i*(width+spacing)
    ybottom<-j*(width+spacing)
    rect(xleft+width/2,
         ybottom,
         xleft+width,
         ybottom+width,
         col=colrange[round(DEM_sd_scaledcard[i,j]*100)],
         border = FALSE)
  }
}
```

Perform PCA on the data, which tells us that there are some funny outliers:
```{r}
summarized<-aggregate(.~joined_d$Xc+joined_d$Yc,joined_d,FUN = mean)
summarized$dE<-0
for(i in 1:dim(summarized)[1]){
  summarized$dE[i]<-deltaE2000(c(summarized$L_measured[i],
                                 summarized$a_measured[i],
                                 summarized$b_measured[i]),
                               c(summarized$L_master[i],
                                 summarized$a_master[i],
                                 summarized$b_master[i]))
}

summarized$dE_scaled<-(summarized$dE-mean(summarized$dE))/sd(summarized$dE)

pca.out<-prcomp(joined_d[c("DE_scaled","Xc","Yc")],scale. = TRUE)
biplot(pca.out, xlabs=rep("·", nrow(joined_d)))
```

Try a different color scheme:
```{r}
xspace<-cbind(1:100,1:100)
plot(xspace,type="n",axes=FALSE,xlab = "",ylab = "")
start<-5
width<-8
spacing<-2

dEpallette <- colorRampPalette(c("green", "red"))(max(ceiling(log(summarized$dE))))

for( i in 1:8){
  for( j in 1:8){
    xleft<-i*(width+spacing)
    ybottom<-j*(width+spacing)
    log_dE<-log(summarized$dE[which(summarized$Xc==i&summarized$Yc==j)])
    rect(xleft,ybottom,xleft+width,ybottom+width,col=dEpallette[ceiling(log_dE)],
         border = FALSE)
  }
}
```

Plot the means and variances of all sheets (right: means, left: variances).
```{r}
par(mfrow=c(1,2))
sheets <- matrix(nrow=13,ncol=2688)
sheetMeans <- c()
sheetVars <- c()

for (i in 1:13) {
  sheets[i,] <- joined_d$DE[which(joined_d$sheet==i)]
  sheetMeans[i] <- mean(joined_d$DE[which(joined_d$sheet==i)])
  sheetVars[i] <- var(joined_d$DE[which(joined_d$sheet==i)])
}

boxplot(sheetMeans)
boxplot(sheetVars)
par(mfrow=c(1,1))
```

Display the master color card (some colors are missing due to errors resulting from converting Lab to RGB):
```{r}
xspace<-cbind(1:100,1:100)
plot(xspace,type="n",axes=FALSE,xlab = "",ylab = "")
start<-5
width<-8
spacing<-2

labVals <- c()
for( i in 1:8){
  for( j in 1:8){
    xleft<-i*(width+spacing)
    ybottom<-j*(width+spacing)
    labVals <- masterCard[(masterCard$Crow==i & masterCard$Ccol==j), ][,c("L", "a", "b")]
    
    b <- convertColor(c(labVals$L, labVals$a, labVals$b),from="Lab",to="Apple RGB",clip=TRUE)
    
    tryCatch({
      cola <- rgb(b[1,][1], b[1,][2], b[1,][3])
      rect(xleft,ybottom,xleft+width,ybottom+width,col=cola,border = FALSE)
    }, warning = function(w) {
    }, error = function(e) {print(c(i,j,b))})
  }
}
```

Which color deviates the most across all sheets from the master (highest delta E)? Top rectangle is what the color looks like in the master sheet. Bottom rectangle is what the average of the most deviated color looks like.
```{r}
maxDeVal <- max(summarized$dE, na.rm = TRUE)
maxCol <- summarized[which.max(summarized$dE),]

xspace<-cbind(1:100,1:100)
plot(xspace,type="n",axes=FALSE,xlab = "",ylab = "")
start<-1
width<-30
spacing<-2

a <- convertColor(c(maxCol$L_master, maxCol$a_master, maxCol$b_master),
                  from="Lab",to="sRGB",clip=NA)
b <- convertColor(c(maxCol$L_measured, maxCol$a_measured, maxCol$b_measured),
                  from="Lab",to="sRGB",clip=NA)
a_color <- rgb(a[1,][1], a[1,][2], a[1,][3])
b_color <-  rgb(b[1,][1], b[1,][2], b[1,][3])

xleft<-1*(width+spacing)
ybottom<-1*(width+spacing)
rect(xleft,ybottom,xleft+width,ybottom+width,col=a_color,border = FALSE)

xleft<-1*(width+spacing)
ybottom<-2*(width+spacing)
rect(xleft,ybottom,xleft+width,ybottom+width,col=b_color,border = FALSE)
```

The delta of the above E colors is `r maxDeVal`, which means (according to [this chart](http://zschuessler.github.io/DeltaE/learn/)) the colors are more similar than opposite. However, the difference is perceptible to the common human eye.

Which color deviates the least (lowest delta E)? Top rectangle is what the color looks like in the master sheet. Bottom rectangle is what the average of the most deviated color looks like.
```{r}
minDeVal <- min(summarized$dE, na.rm = TRUE)
minCol <- summarized[which.min(summarized$dE),]

xspace<-cbind(1:100,1:100)
plot(xspace,type="n",axes=FALSE,xlab = "",ylab = "")
start<-1
width<-30
spacing<-2

a <- convertColor(c(minCol$L_master, minCol$a_master, minCol$b_master),
                  from="Lab",to="sRGB",clip=NA)
b <- convertColor(c(minCol$L_measured, minCol$a_measured, minCol$b_measured),
                  from="Lab",to="sRGB",clip=NA)
a_color <- rgb(a[1,][1], a[1,][2], a[1,][3])
b_color <-  rgb(b[1,][1], b[1,][2], b[1,][3])

xleft<-1*(width+spacing)
ybottom<-1*(width+spacing)
rect(xleft,ybottom,xleft+width,ybottom+width,col=a_color,border = FALSE)

xleft<-1*(width+spacing)
ybottom<-2*(width+spacing)
rect(xleft,ybottom,xleft+width,ybottom+width,col=b_color,border = FALSE)
```
Building the mean of L,a,b of all colorcards, convert it to RGB and compare it with the mastercard.Again you can see only one strong outlier.
```{r}
Xc <- rep(1:8, each =8)
Yc <- rep(1:8,8)
Labavg <- data.frame(Xc,Yc)
L <- zeros(8*8,1)
a <- zeros(8*8,1)
b <- zeros(8*8,1)

for( i in 1:8){
  row <- joined_d[joined_d$Xc == i,]
  for( j in 1:8){
    color <- row[row$Yc == j,]
    L[(i-1)*8+j] <- mean(color$L_measured)
    a[(i-1)*8+j] <- mean(color$a_measured)
    b[(i-1)*8+j] <- mean(color$b_measured)
  }
}

Labavg$L <- L
Labavg$a <- a
Labavg$b <- b


xspace<-cbind(1:100,1:100)
plot(xspace,type="n",axes=FALSE,xlab = "",ylab = "")
start<-5
width<-8
spacing<-2

labVals <- c()
for( i in 1:8){
  for( j in 1:8){
    xleft<-i*(width+spacing)
    ybottom<-j*(width+spacing)
    labVals <- Labavg[(Labavg$Xc==i & Labavg$Yc==j), ][,c("L", "a", "b")]
    
    b <- convertColor(c(labVals$L, labVals$a, labVals$b),from="Lab",to="Apple RGB",clip=TRUE)
    
    tryCatch({
      cola <- rgb(b[1,][1], b[1,][2], b[1,][3])
      rect(xleft,ybottom,xleft+width,ybottom+width,col=cola,border = FALSE)
    }, warning = function(w) {
    }, error = function(e) {print(c(i,j,b))})
  }
}
labVals <- c()
for( i in 1:8){
  for( j in 1:8){
    xleft<-i*(width+spacing)
    ybottom<-j*(width+spacing)
    labVals <- masterCard[(masterCard$Crow==i & masterCard$Ccol==j), ][,c("L", "a", "b")]
    
    b <- convertColor(c(labVals$L, labVals$a, labVals$b),from="Lab",to="Apple RGB",clip=TRUE)
    
    tryCatch({
      cola <- rgb(b[1,][1], b[1,][2], b[1,][3])
      rect(xleft,ybottom,xleft+width/2,ybottom+width,col=cola,border = FALSE)
    }, warning = function(w) {
    }, error = function(e) {print(c(i,j,b))})
  }
}

```
The delta of the above E colors is `r minDeVal`, which means (according to [this chart](http://zschuessler.github.io/DeltaE/learn/)) that the difference between the colors is not perceptible to the common human eye.

Heat map of the summarized Lab values across all sheets against their master values:
```{r}
x <- summarized[,c("L_measured", "a_measured", "b_measured",
                   "L_master", "a_master", "b_master")]

# Use 'scale' to normalize (right)
heatmap(as.matrix(x), scale="column", col = terrain.colors(256), margins=c(8,5))
```
